\vskip-0.6in Endogenous Politics and the Design of Trade Institutions

### 2015-01-08 ex_ante.tex
convert lines 52-60 ex_ante to text

Thus I model the governments ex-ante preferences as ... threshold

	\item Thus the weight the gov't places on the lobbying industry's profits will change if ex-ante effort is high enough. The trade agreement tariff cap will be set according to $\ga(e^{EA})$ instead of at $\tau^R_{W,e}$ if that would lead to a higher trade agreement tariff (if $\tau = \frac{4(\ga(e^{EA}-1)}{25-4\ga(e^{EA}} < \tau^R_{W,e}$, then government prefers to set ex-post tariff)
			\begin{itemize}
				\item Recall how $\tau^{TA}$ is determined under ex-post lobbying only. Because the trade agreement acts as a commitment device, the government chooses the tariff level $\tau^R_{W,e}$, and implicitly $e^{EP}=e^R_{W,e}$, to maximize the joint political welfare of the two governments as in Equation~\ref{exp:1}.
		\begin{itemize}
			\item Government DOESN'T have to just set trade agreement tariff given what it expects will happen in the future; it can influence lobby's future behavior through the trade agreement (at least on the high side; it can't encourage extra lobbying, which it would like to do in some cases, e.g. exact tariff commitment)
			\item With ex-post lobbying only, at ex-ante stage gov't has to take into account how ex-post lobbying will affect it; picks $\tau^{TA}$ to max joint political welfare subject to this constraint (sets tariff to produce optimal $\ga$ according to $\tau^N$)
		\end{itemize}
				\item This is where gov't loses some measure of control with ex-ante lobbying: 
					\begin{itemize}
						\item If $e^{EA}$ is low enough, ex-ante lobbying has no impact. Government behaves as in ex-post case and chooses $\tau^R_{W,e}$.
						\item If $e^{EA}$ is above some threshold, $\tau^{TA} >\tau^{TA^*}$ b/c $e^{EA}$ now determines $\tau^{TA}$
						\item Gov't still chooses ``optimal'' trade agreement, but it's constrained optimal. Constrained by current pressure.
					\end{itemize}		
				\end{itemize}
	\item Real question: why does the government set non-zero $\tau^R_{W,e}$?
		\begin{itemize}
			\item Because joint welfare is higher when $e^R_{W,e}$ together with $\tau^R_{W,e}$ than it is when $e=0$ and $\tau=0$. These are fundamentally politically motivated governments.
				\begin{itemize}
					\item But depending on how welfare changes as pressure increases, gov't may be looking ahead and manipulating lobbying incentives as shown in Section~\ref{sec:objfcn}
					\item When ex-ante lobbying is possible, in order to respond to that pressure, govts may have to raise $\tau^{TA}$ above $\tau^R_{W,e}$ (lobby won't ever complain if tariff cap is higher than what it asks for)
					\item That is modeled here with a simple cut-off: if $e^{EA}$ is larger than anticipated $e^{EP}$, gov't responds to $e^{EA}$ instead of using trade agreementto influence $e^{EP}$
				\end{itemize}
			\item We will see that this change to the government's preferences has the impacts one would expect, at least in the case of tariff caps: net profits for the lobby (weakly) increase, and net political utility for the government (weakly) decreases
		\end{itemize}
\end{itemize}

\newpage
Exact tariff commitment result
\begin{itemize}
	\item We know $e^{EP} =0$ and $\tau^{EP} = \tau^R_{S,e}$ when no possibility of ex-ante lobbying
	\item What does lobby's problem look like?
		\begin{itemize}
			\item At ex-post stage, $e^{EA}$ sunk. So ex-post problem looks the same.
			\item In principle, at ex-ante stage, $e^{EA}$ will have impact on $e^{EP}$. But not in exact tariff case: $e^{EP} = 0 \ \forall e^{EP}$.
			\item So this collapses to a simple problem of maximizing joint gov't welfare with only $e=e^{EA}$
				\begin{itemize}
					\item Looks like a problem with no ex-post lobbying, but where there is perfect enforcement of the tariff commitment
					\item Ex-post, gov't will have $\ga(0)$ but $\tau^A = \tau^R_{S,e}$. Gov't would prefer $\tau^A = 0$, so this is a welfare loss relative to no ex-ante case at this stage.
						\begin{itemize}
							\item Perhaps want note relating to MRC here?
						\end{itemize}
					\item Without knowing how gov't welfare varies in $e$ (and therefore $\ga$), we can't say for sure the effect at the ex-ante stage. As with example in Figure~\ref{fig:weight}, welfare could be reduced (certainly will for foreign), but could also be increased. Again, this is a kind of political welfare from negotiations, not an expected real welfare, so not clear how to weigh it against loss in ex-post stage that is a real loss.
					\item Very clear gain for lobby. If there were no gain, it could just choose $e^{EA} = 0$
				\end{itemize}
		\end{itemize}
\end{itemize}

>>>>

Thus I model the influence of ex-ante lobbying on the governments' tariff-setting behavior as operating through a threshold. Recall how the trade agreement tariff is determined under ex-post lobbying only. Because the trade agreement acts as a commitment device, the governments choose the tariff level $\tau^R_{W,e}$, and implicitly $e^{EP}=e^R_{W,e}$, to maximize their joint political welfare as in Equation~\ref{exp:1}, subject only to the constraint that $\ga$ will be realized according to the ex-post lobbying process.

As discussed in Section~\ref{sec:objfcn}, depending on how government welfare changes as political pressure increases, the governments may be looking ahead and manipulating ex-post lobbying incentives when there is only ex-post lobbying. When ex-ante lobbying is possible the governments may have to raise the trade agreement tariff above $\tau^R_{W,e}$ in order to respond to that pressure.

This is modeled here with a simple cut-off. The trade agreement tariff cap will be set according to $\ga\left(e^{EA}\right)$ instead of at $\tau^R_{W,e}$ if $\ga\left(e^{EA}\right)$ would lead to a higher trade agreement tariff. If, on the other hand, $\tau = \frac{4(\ga(e^{EA})-1)}{25-4\ga(e^{EA})} < \tau^R_{W,e}$, the governments respond to the higher ex-post pressure and set $\tau^R_{W,e}$.

Thus the governments lose some measure of control with ex-ante lobbying. If $e^{EA}$ is low enough, ex-ante lobbying has no impact. But if $e^{EA}$ is above the threshold, the trade agreement tariff is strictly greater than $\tau^R_{W,e}$ and so joint government welfare will not be at its globally optimal level. The trade agreement the governments choose will be constrained by the ex-ante pressure.

We will see that this change to the government's preferences has the impacts one would expect, at least in the case of tariff caps: net profits for the lobby (weakly) increase, and net political utility for the governments (weakly) decrease.

Exact tariff commitment result
\begin{itemize}
	\item We know $e^{EP} =0$ and $\tau^{EP} = \tau^R_{S,e}$ when no possibility of ex-ante lobbying
	\item What does lobby's problem look like?
		\begin{itemize}
			\item At ex-post stage, $e^{EA}$ sunk. So ex-post problem looks the same.
			\item In principle, at ex-ante stage, $e^{EA}$ will have impact on $e^{EP}$. But not in exact tariff case: $e^{EP} = 0 \ \forall e^{EP}$.
			\item So this collapses to a simple problem of maximizing joint gov't welfare with only $e=e^{EA}$
				\begin{itemize}
					\item Looks like a problem with no ex-post lobbying, but where there is perfect enforcement of the tariff commitment
					\item Ex-post, gov't will have $\ga(0)$ but $\tau^A = \tau^R_{S,e}$. Gov't would prefer $\tau^A = 0$, so this is a welfare loss relative to no ex-ante case at this stage.
						\begin{itemize}
							\item Perhaps want note relating to MRC here?
						\end{itemize}
					\item Without knowing how gov't welfare varies in $e$ (and therefore $\ga$), we can't say for sure the effect at the ex-ante stage. As with example in Figure~\ref{fig:weight}, welfare could be reduced (certainly will for foreign), but could also be increased. Again, this is a kind of political welfare from negotiations, not an expected real welfare, so not clear how to weigh it against loss in ex-post stage that is a real loss.
					\item Very clear gain for lobby. If there were no gain, it could just choose $e^{EA} = 0$
				\end{itemize}
		\end{itemize}
\end{itemize}
